---
title: "Zenith Abacus Estimation"
author: "JS; template for RMD from mosaic; lots of code from MCF"
date: "June 1, 2015"
output: 
  html_document:
    fig_height: 6
    fig_width: 6
  pdf_document:
    fig_height: 7
    fig_width: 7
  word_document:
    fig_height: 7
    fig_width: 7
---

```{r include=FALSE}
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r, include=FALSE}
# Load additional packages here.  Uncomment the line below to use Project MOSAIC data sets.
# require(mosaicData)   
```

Load more libraries
```{r, include = FALSE}
library(plyr)
library(reshape2)
library(ggplot2)
library(lme4)
rm(list=ls())
theme_set(theme_bw())
options(warn=-1)
```


Read in the data first
```{r}
d <- read.csv("/Users/jsulliv1/Desktop/UCSDProjects/AnalyzeWrite/Abacus/zenith all data complete and correct estimation 040815.csv")
d$condition <- factor(d$condition, levels=c("control","abacus"))
head(d)

```

Combine Woodcock and Wiat and z-score the Standardized Test composite
```{r}
d$standardized <- rowMeans(d[,c("wiat","woodcock")],na.rm=TRUE) 

summary(d$standardized[d$year==0])
summary(d$standardized[d$year==1])
summary(d$standardized[d$year==2])
summary(d$standardized[d$year==3])
summary(d$standardized[d$condition =="abacus"])
summary(d$standardized[d$condition =="control"])


```

Descriptives of Each variable
```{r}
summary(d$ans[d$condition =="abacus"])
summary(d$ans[d$condition =="control"])
summary(d$ans[d$year==0])
summary(d$ans[d$year==1])
summary(d$ans[d$year==2])
summary(d$ans[d$year==3])





summary(d$deviance[d$condition =="abacus"])
summary(d$deviance[d$condition =="control"])
summary(d$deviance[d$year==1])
summary(d$deviance[d$year==2])
summary(d$deviance[d$year==3])




summary(d$linr2[d$condition =="abacus"])
summary(d$linr2[d$condition =="control"])
summary(d$linr2[d$year==1])
summary(d$linr2[d$year==2])
summary(d$linr2[d$year==3])


summary(d$ordinality[d$condition =="abacus"])
summary(d$ordinality[d$condition =="control"])
summary(d$ordinality[d$year==1])
summary(d$ordinality[d$year==2])
summary(d$ordinality[d$year==3])

```



Reliability for Estimation tasks
```{r}
deviance.corrs <- data.frame(year=factor(c("1-2","2-3")),
                             corrs=c(cor.test(d$deviance[d$year==1],d$deviance[d$year==2])$estimate,
                                     cor.test(d$deviance[d$year==2],d$deviance[d$year==3])$estimate))
deviance.corrs


linr2.corrs <- data.frame(year=factor(c("1-2","2-3")),
                          corrs=c(cor.test(d$linr2[d$year==1],d$linr2[d$year==2])$estimate,
                                  cor.test(d$linr2[d$year==2],d$linr2[d$year==3])$estimate))
linr2.corrs


ord.corrs <- data.frame(year=factor(c("1-2","2-3")),
                          corrs=c(cor.test(d$ordinality[d$year==1],d$ordinality[d$year==2])$estimate,
                                  cor.test(d$ordinality[d$year==2],d$ordinality[d$year==3])$estimate))
ord.corrs


ans.corrs <- data.frame(year=factor(c("0-1","1-2","2-3")),
                        corrs=c(cor.test(d$ans[d$year==0],d$ans[d$year==1])$estimate, cor.test(d$ans[d$year==1],d$ans[d$year==2])$estimate,
                                cor.test(d$ans[d$year==2],d$ans[d$year==3])$estimate))
ans.corrs

```
We see that the correlations are pretty decent year-to-year for our three DVs.


OK, now we test for effect of training for ANS. We can use a model b/c we have all 4 years:
```{r}
w.interaction <- lmer(ans ~ year  * condition + (subnum|year), data=d)
summary(w.interaction)
wo.interaction<-lmer(ans ~ year + condition + (subnum|year), data=d)
a<-anova(w.interaction, wo.interaction)
a

```



Now, we do t-tests by task to find differences between control and abacus groups. We will correct for multiple comparisons post-hoc, correcting for the # of comparisons per DV. 

```{r}
tasks <- c("deviance","linr2","ans", "ordinality")

for (t in tasks) {
  md <- melt(d, id.vars=c("subnum","year","condition","female","age"), measure.vars=t)
  
  
  print(paste("*********",t,"**********"))
  print(t.test(subset(md,condition=="abacus" & year==1)$value,
               subset(md,condition=="control" & year==1)$value,var.equal=TRUE))
  print(t.test(subset(md,condition=="abacus" & year==2)$value,
               subset(md,condition=="control" & year==2)$value,var.equal=TRUE))
  print(t.test(subset(md,condition=="abacus" & year==3)$value,
               subset(md,condition=="control" & year==3)$value,var.equal=TRUE))
}
```


PLOTS
ANS
Without intervention split

```{r, warnings=FALSE, message=FALSE}
qplot(ans, standardized, facets=~year, 
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Weber Fraction")        
```

And with:
  
  ```{r, warnings=FALSE, message=FALSE}
qplot(ans, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Weber Fraction")        
```

PAE
Now do the same for deviance:
  
  ```{r, warnings=FALSE}
qplot(deviance, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("PAE")    

qplot(deviance, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("PAE")    

```

and for linear r^2:
  
  ```{r, warnings=FALSE}
qplot(linr2, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Linear r^2")    

qplot(linr2, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Linear r^2")       
```

and for ordinality:
  
  ```{r, warnings=FALSE}
qplot(ordinality, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Ordinality")    

qplot(ordinality, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Ordinality")       
```




Setup Models
```{r, echo=FALSE}
model.pred <- function (data, 
                        mathvars=c("wiat","woodcock","placeval","math","arith","standardized"),
                        predvars=c("ans","mental.rot","verbalwm",
                                   "spatialwm","ravens","deviance","linr2","ordinality"),
                        model.formula, 
                        model.name="model",
                        years = c(0,1,2,3)) {  
  
  coefs <- data.frame()
  for (yr in years) {
    ## Ensure complete cases for each year
    y <- subset(data, year==yr)
    yc <- y[complete.cases(y[, mathvars]),]
    controls <- subset(yc, condition == "control")
    
    ## standardize predictors
    for (v in c(predvars,mathvars)) {
      eval(parse(text=paste("yc$",v,".scale <- scale(yc$",v,")",sep="")))
    }
    
    ## Find principal components
    pca <- princomp(~wiat.scale + woodcock.scale + math.scale + 
                      placeval.scale + arith.scale, 
                    data=yc, 
                    cor = TRUE, scores = TRUE)  
    yc$pc1 <- pca$scores[,"Comp.1"]
    
    ## run model
    model <- lm(model.formula, data=yc)  
    coefs <- rbind.fill(coefs, 
                        data.frame(year=yr,
                                   betas=coef(model),
                                   se=coef(summary(model))[, "Std. Error"],
                                   preds=names(coef(model)),
                                   model=model.name))
  }

  return(coefs)
}
```

And a function to relevel coefficients. (Also not rendered). 

```{r, echo=FALSE}
relevel.coefs <- function (coefs, dv="ans.scale") {
  
  coefs$preds <- factor(coefs$preds, 
                        levels=c("(Intercept)",dv,"mental.rot.scale",
                                 "spatialwm.scale","verbalwm.scale","ravens.scale",
                                 "age","conditionabacus"))
  coefs$preds <- revalue(coefs$preds, c("(Intercept)"="Intercept",
                                        "ans.scale" ="ANS",
                                        "linr2.scale"="Estimation Linear r^2",
                                        "ordinality.scale"="Ordinality",
                                        "deviance.scale"="PAE",
                                        "mental.rot.scale"="Mental Rotation",
                                        "ravens.scale"="Raven's",
                                        "spatialwm.scale"="Spatial WM",
                                        "verbalwm.scale"="Verbal WM",
                                        "conditionabacus"="Intervention",
                                        "age"="Age"))
  coefs$model <- factor(coefs$model, levels=c("vanilla","complex","complex + cond"))
  coefs$model <- revalue(coefs$model, c("vanilla"="Univariate",
                                        "complex"="All Preds",
                                        "complex + cond"=
                                          "All Preds + Intervention"))
  return(coefs)
}
```

Actually Test Models

ANS and standardized test composite

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ ans.scale), 
                              model.name="vanilla")),
            
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ ans.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond")))

coefs <- relevel.coefs(coefs)


qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 

```

Percent Absolute Error

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        deviance.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        deviance.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="deviance.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
```

Linear R2

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        linr2.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        linr2.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="linr2.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
```


Ordinality

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ordinality.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="ordinality.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
```


Principle Components:
Here we take a principal components analysis of all the math measures and then take PC1, which is the factor with the largest explained variance amongst all the math measures. This is just one way of making a composite of all the math DVs and exploring its relatedness to ANS. Note: the PCs come out negative, so MCF flipped them for interpretability.


ANS
```{r}
```{r}
coefs <- rbind(model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale), 
                          model.name="vanilla"),
               model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale + 
                                                    mental.rot.scale + 
                                                    verbalwm.scale + 
                                                    spatialwm.scale + ravens.scale + 
                                                    age), 
                          model.name="complex"),
               model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale + 
                                                    mental.rot.scale + 
                                                    verbalwm.scale + 
                                                    spatialwm.scale + ravens.scale + 
                                                    age + condition), 
                          model.name="complex + cond"))
coefs <- relevel.coefs(coefs)

qplot(preds,-betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=-betas-se,
                     ymax=-betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
```
Ordinality
```{r}

coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ordinality.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="ordinality.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
```

PAE

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        deviance.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        deviance.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="deviance.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 

```


LinR2


```{r}

coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        linr2.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        linr2.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="linr2.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
  
```


Now, combine all estimation measures
Predicting PC1

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ans.scale + deviance.scale + linr2.scale + ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ans.scale + 
                                                        deviance.scale + 
                                                        linr2.scale + 
                                                        ordinality.scale +
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv=c("ans.scale", "deviance.scale",
                                   "linr2.scale","ordinality.scale"))

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 

```


Now, combine all estimation measures
Predicting standardized test

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale + deviance.scale + linr2.scale + ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale + 
                                                        deviance.scale + 
                                                        linr2.scale + 
                                                        ordinality.scale +
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv=c("ans.scale", "deviance.scale",
                                   "linr2.scale","ordinality.scale"))

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
```

STILL NEED: Figure that Mike will make.




OK, so now that we have that pretty stuff, it's time to do model comparisons that Jess knows how to get the betas out of. 

Univariate analyses

Here are all the correlations:

```{r}
ddply(d, .(year), 
      summarise, 
      ans.std.cor = try(cor.test(ans,standardized)$estimate),
      ans.pc1.cor = try(cor.test(ans,pc1)$estimate),
      dev.pc1.cor = try(cor.test(deviance,pc1)$estimate),
      dev.std.cor = try(cor.test(deviance,standardized)$estimate),
      lin.pc1.cor = try(cor.test(linr2,pc1)$estimate),
      lin.std.cor = try(cor.test(linr2,standardized)$estimate),
      ord.pc1.cor = try(cor.test(ordinality,pc1)$estimate),
      ord.std.cor = try(cor.test(ordinality,standardized)$estimate))
```

And their p-values:
```{r}
ddply(d, .(year), 
      summarise, 
      ans.std.cor = try(cor.test(ans,standardized)$p.value),
      ans.pc1.cor = try(cor.test(ans,pc1)$p.value),
      dev.pc1.cor = try(cor.test(deviance,pc1)$p.value),
      dev.std.cor = try(cor.test(deviance,standardized)$p.value),
      lin.pc1.cor = try(cor.test(linr2,pc1)$p.value),
      lin.std.cor = try(cor.test(linr2,standardized)$p.value),
      ord.pc1.cor = try(cor.test(ordinality,pc1)$p.value),
      ord.std.cor = try(cor.test(ordinality,standardized)$p.value))
```


