---
title: "Zenith Abacus Estimation"
author: "JS; template for RMD from mosaic; lots of code from MCF"
date: "June 3, 2015"
output: 
  html_document:
    fig_height: 8
    fig_width: 8
  pdf_document:
    fig_height: 7
    fig_width: 7
  word_document:
    fig_height: 7
    fig_width: 7
---

```{r include=FALSE}
# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  cache=TRUE,
  size="small"    # slightly smaller font for code
)
```

Load more libraries
```{r, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
# require(mosaic)
rm(list=ls())
theme_set(theme_bw())
options(warn=-1)
```

Read in the data first
```{r}
d <- read.csv("zenith all data.csv")
d$condition <- factor(d$condition, levels=c("control","abacus"))
head(d)
```

Combine Woodcock and Wiat and z-score the Standardized Test composite
```{r}
d$standardized <- rowMeans(d[,c("wiat","woodcock")],na.rm=TRUE) 

summary(d$standardized[d$year==0])
summary(d$standardized[d$year==1])
summary(d$standardized[d$year==2])
summary(d$standardized[d$year==3])
summary(d$standardized[d$condition =="abacus"])
summary(d$standardized[d$condition =="control"])
```

# Descriptives and preliminaries

```{r}
summary(d$ans[d$condition =="abacus"])
summary(d$ans[d$condition =="control"])
summary(d$ans[d$year==0])
summary(d$ans[d$year==1])
summary(d$ans[d$year==2])
summary(d$ans[d$year==3])

summary(d$deviance[d$condition =="abacus"])
summary(d$deviance[d$condition =="control"])
summary(d$deviance[d$year==1])
summary(d$deviance[d$year==2])
summary(d$deviance[d$year==3])

summary(d$linr2[d$condition =="abacus"])
summary(d$linr2[d$condition =="control"])
summary(d$linr2[d$year==1])
summary(d$linr2[d$year==2])
summary(d$linr2[d$year==3])

summary(d$ordinality[d$condition =="abacus"])
summary(d$ordinality[d$condition =="control"])
summary(d$ordinality[d$year==1])
summary(d$ordinality[d$year==2])
summary(d$ordinality[d$year==3])

```

## Reliability for Estimation tasks

Deviance.
```{r}
data.frame(year=factor(c("1-2","2-3")),
           corrs=c(cor.test(d$deviance[d$year==1],
                            d$deviance[d$year==2])$estimate,
                   cor.test(d$deviance[d$year==2],
                            d$deviance[d$year==3])$estimate))
```

Linear $r^2$.
```{r}
data.frame(year=factor(c("1-2","2-3")),
           corrs=c(cor.test(d$linr2[d$year==1],d$linr2[d$year==2])$estimate,
                   cor.test(d$linr2[d$year==2],d$linr2[d$year==3])$estimate))
```

Ordinality.
```{r}
data.frame(year=factor(c("1-2","2-3")),
           corrs=c(cor.test(d$ordinality[d$year==1],
                            d$ordinality[d$year==2])$estimate,
                   cor.test(d$ordinality[d$year==2],
                            d$ordinality[d$year==3])$estimate))
```

ANS.
```{r}
data.frame(year=factor(c("0-1","1-2","2-3")),
           corrs=c(cor.test(d$ans[d$year==0],
                            d$ans[d$year==1])$estimate, 
                   cor.test(d$ans[d$year==1],
                            d$ans[d$year==2])$estimate,
                   cor.test(d$ans[d$year==2],
                            d$ans[d$year==3])$estimate))
```
We see that the correlations are pretty decent year-to-year for our three DVs.

## Training effects

OK, now we test for effect of training for ANS. We can use a model b/c we have all 4 years:
```{r}
w.interaction <- lmer(ans ~ year  * condition + (subnum|year), data=d)
summary(w.interaction)
wo.interaction<-lmer(ans ~ year + condition + (subnum|year), data=d)
a<-anova(w.interaction, wo.interaction)
a

```

Now, we do t-tests by task to find differences between control and abacus groups. We will correct for multiple comparisons post-hoc, correcting for the # of comparisons per DV. 

```{r}
tasks <- c("deviance","linr2","ans", "ordinality")

for (t in tasks) {
  md <- d %>% 
    rename_("value" = t)
  
  print(paste("*********",t,"**********"))
  print(t.test(subset(md,condition=="abacus" & year==1)$value,
               subset(md,condition=="control" & year==1)$value,var.equal=TRUE))
  print(t.test(subset(md,condition=="abacus" & year==2)$value,
               subset(md,condition=="control" & year==2)$value,var.equal=TRUE))
  print(t.test(subset(md,condition=="abacus" & year==3)$value,
               subset(md,condition=="control" & year==3)$value,var.equal=TRUE))
}
```

# PLOTS

ANS - Without intervention split
```{r, warnings=FALSE, message=FALSE}
qplot(ans, standardized, facets=~year, 
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Weber Fraction")        
```

And with:
```{r, warnings=FALSE, message=FALSE}
qplot(ans, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Weber Fraction")        
```

Now do the same for deviance:
```{r, warnings=FALSE}
qplot(deviance, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("PAE")    

qplot(deviance, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("PAE")    

```

and for linear $r^2$:
```{r, warnings=FALSE}
qplot(linr2, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Linear r^2")    

qplot(linr2, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Linear r^2")       
```

and for ordinality:
```{r, warnings=FALSE}
qplot(ordinality, standardized, facets=~year,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Ordinality")    

qplot(ordinality, standardized, facets=~year, 
      col=condition,
      data=d) + 
  geom_smooth(method="lm") +
  ylab("Standardized Test Composite") +
  xlab("Ordinality")       
```

# Models

Set up models - doesn't currently show. 
```{r, echo=FALSE}
model.pred <- function (data, 
                        mathvars=c("wiat","woodcock","placeval",
                                   "math","arith","standardized"),
                        predvars=c("ans","mental.rot","verbalwm",
                                   "spatialwm","ravens","deviance",
                                   "linr2","ordinality"),
                        model.formula, 
                        model.name="model",
                        years = c(0,1,2,3)) {  
  
  coefs <- data.frame()
  for (yr in years) {
    ## Ensure complete cases for each year
    y <- subset(data, year==yr)
    yc <- y[complete.cases(y[, mathvars]),]
    controls <- subset(yc, condition == "control")
    
    ## standardize predictors
    for (v in c(predvars,mathvars)) {
      eval(parse(text=paste("yc$",v,".scale <- scale(yc$",v,")",sep="")))
    }
    
    ## Find principal components
    pca <- princomp(~wiat.scale + woodcock.scale + math.scale + 
                      placeval.scale + arith.scale, 
                    data=yc, 
                    cor = TRUE, scores = TRUE)  
    yc$pc1 <- pca$scores[,"Comp.1"]
    
    ## run model
    model <- lm(model.formula, data=yc)  
    coefs <- bind_rows(coefs, 
                       data.frame(year=yr,
                                  betas=coef(model),
                                  se=coef(summary(model))[, "Std. Error"],
                                  preds=names(coef(model)),
                                  model=model.name))
    }
  
  return(coefs)
}
```

And a function to relevel coefficients. (Also not rendered). 
```{r, echo=FALSE}
relevel.coefs <- function (coefs, dv="ans.scale") {
  
  coefs$preds <- factor(coefs$preds, 
                        levels=c("(Intercept)",dv,"mental.rot.scale",
                                 "spatialwm.scale","verbalwm.scale","ravens.scale",
                                 "age","conditionabacus"))
  coefs$preds <- plyr::revalue(coefs$preds, c("(Intercept)"="Intercept",
                                        "ans.scale" ="ANS",
                                        "linr2.scale"="Estimation Linear r^2",
                                        "ordinality.scale"="Ordinality",
                                        "deviance.scale"="PAE",
                                        "mental.rot.scale"="Mental Rotation",
                                        "ravens.scale"="Raven's",
                                        "spatialwm.scale"="Spatial WM",
                                        "verbalwm.scale"="Verbal WM",
                                        "conditionabacus"="Intervention",
                                        "age"="Age"))
  coefs$model <- factor(coefs$model, levels=c("vanilla","complex","complex + cond"))
  coefs$model <- plyr::revalue(coefs$model, c("vanilla"="Univariate",
                                        "complex"="All Preds",
                                        "complex + cond"=
                                          "All Preds + Intervention"))
  return(coefs)
}
```

## Test Models on standardized tests

ANS and standardized test composite
```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale), 
                              model.name="vanilla")),               
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond")))
coefs <- relevel.coefs(coefs)

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
ans.coefs <- coefs
as.data.frame(coefs)
```

Percent Absolute Error
```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        deviance.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        deviance.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="deviance.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
pae.coefs <- coefs
as.data.frame(coefs)
```

Linear $r^2$
```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        linr2.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        linr2.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="linr2.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
as.data.frame(coefs)
linr2.coefs <- coefs
```

Ordinality
```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ordinality.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="ordinality.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
ord.coefs <- coefs
as.data.frame(coefs)
```

Now for the guts of one of the models, just an example from year 0, ANS. This uses ANOVA for model comparison. Section not rendered since I don't think we'll use it; Jess is going to try to address github scaling issue.
```{r}
mathvars=c("wiat","woodcock","placeval","math","arith","standardized")
predvars=c("ans","mental.rot","verbalwm","spatialwm","ravens","deviance",
           "linr2","ordinality")

y <- subset(d, year==0)
yc <- y[complete.cases(y[, c("ans","mental.rot","verbalwm","spatialwm","ravens")]),]
    
## standardize predictors in
for (v in c(predvars,mathvars)) {
  eval(parse(text=paste("yc$",v,".scale <- scale(yc$",v,")",sep="")))
}
        
## run models
model1 <- lm(standardized.scale ~ ans.scale + mental.rot.scale + verbalwm.scale + 
              spatialwm.scale + ravens.scale +  age + condition, data = yc)
model2 <- lm(standardized.scale ~ mental.rot.scale + verbalwm.scale + 
              spatialwm.scale + ravens.scale +  age + condition, data = yc)
summary(model1)
summary(model2)
anova(model1, model2)
```

Year 0 Analyses by JS, to check betas
```{r}
#ensure year 0, ensure complete cases
y0 <- subset(d, year==0)
yc0 <- y0[complete.cases(y0[, c("ans","mental.rot","verbalwm","spatialwm","ravens")]),]


#scale all variables
yc0$standardized.scale <-scale(yc0$standardized)
yc0$ans.scale <-scale(yc0$ans)
yc0$mental.rot.scale<-scale(yc0$mental.rot)
yc0$age.scale<-scale(yc0$age)
yc0$verbalwm.scale<-scale(yc0$verbalwm)
yc0$spatialwm.scale<-scale(yc0$spatialwm)
yc0$ravens.scale<-scale(yc0$ravens)

#ANS, Year 0
withansyear0 <- lm(standardized.scale ~ ans.scale + mental.rot.scale + verbalwm.scale + 
              spatialwm.scale + ravens.scale +  age + condition, data = yc0)
withoutansyear0 <- lm(standardized.scale ~ mental.rot.scale + verbalwm.scale + 
              spatialwm.scale + ravens.scale +  age + condition, data = yc0)
summary(withansyear0)
summary(withoutansyear0)
anova(withansyear0, withoutansyear0)
```
    

## Principal Components
Here we take a principal components analysis of all the math measures and then take PC1, which is the factor with the largest explained variance amongst all the math measures. This is just one way of making a composite of all the math DVs and exploring its relatedness to ANS. Note: the PCs come out negative, so MCF flipped them for interpretability.

ANS
```{r}
coefs <- rbind(model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale), 
                          model.name="vanilla"),
               model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale + 
                                                    mental.rot.scale + 
                                                    verbalwm.scale + 
                                                    spatialwm.scale + ravens.scale + 
                                                    age), 
                          model.name="complex"),
               model.pred(data=d, 
                          model.formula = formula(pc1 ~ ans.scale + 
                                                    mental.rot.scale + 
                                                    verbalwm.scale + 
                                                    spatialwm.scale + ravens.scale + 
                                                    age + condition), 
                          model.name="complex + cond"))
coefs <- relevel.coefs(coefs)

qplot(preds,-betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=-betas-se,
                     ymax=-betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
pca.ans.coefs <- coefs
as.data.frame(coefs)
```

Ordinality
```{r}

coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ordinality.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="ordinality.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
pca.ord.coefs <- coefs
as.data.frame(coefs)
```

PAE
```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        deviance.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        deviance.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="deviance.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
pca.pae.coefs <- coefs
as.data.frame(coefs)
```


Linear $r^2$
```{r}

coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        linr2.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        linr2.scale + 
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv="linr2.scale")

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1))
pca.linr2.coefs <- coefs
as.data.frame(coefs)
```


## Combine all estimation measures
Predicting PC1

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ans.scale + deviance.scale + linr2.scale + ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(pc1 ~ 
                                                        ans.scale + 
                                                        deviance.scale + 
                                                        linr2.scale + 
                                                        ordinality.scale +
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv=c("ans.scale", "deviance.scale",
                                   "linr2.scale","ordinality.scale"))

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 
pca.all.coefs <- coefs
as.data.frame(coefs)
```


Now, combine all estimation measures
Predicting standardized test

```{r}
coefs <- rbind(try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale + deviance.scale + linr2.scale + ordinality.scale), 
                              model.name="vanilla", years=c(1,2,3))),
               try(model.pred(data=d, 
                              model.formula = formula(standardized.scale ~ 
                                                        ans.scale + 
                                                        deviance.scale + 
                                                        linr2.scale + 
                                                        ordinality.scale +
                                                        mental.rot.scale + 
                                                        verbalwm.scale + 
                                                        spatialwm.scale + 
                                                        ravens.scale + 
                                                        age + condition), 
                              model.name="complex + cond", years=c(1,2,3))))
coefs <- relevel.coefs(coefs,dv=c("ans.scale", "deviance.scale",
                                   "linr2.scale","ordinality.scale"))

qplot(preds,betas,facets=model~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust = 1)) 

stand.all.coefs <- coefs
as.data.frame(coefs)
```



```{r, echo=FALSE}
ans.coefs$measure <- "ANS"
ans.coefs$outcome <- "Standarized"
linr2.coefs$measure <- "Linear r^2"
linr2.coefs$outcome <- "Standardized"
ord.coefs$measure <- "Ordinality"
ord.coefs$outcome <- "Standardized"
pae.coefs$measure <- "PAE"
pae.coefs$outcome <- "Standardized"
pca.ans.coefs$measure <- "ANS"
pca.ans.coefs$outcome <- "PCA"
pca.linr2.coefs$measure <- "Linear r^2"
pca.linr2.coefs$outcome <- "PCA"
pca.ord.coefs$measure <- "Ordinality"
pca.ord.coefs$outcome <- "PCA"
pca.pae.coefs$measure <- "PAE"
pca.pae.coefs$outcome <- "PCA"

all.coefs <- bind_rows(ans.coefs, 
                       linr2.coefs, 
                       ord.coefs,
                       pae.coefs,
                       pca.ans.coefs,
                       pca.linr2.coefs, 
                       pca.ord.coefs,
                       pca.pae.coefs) %>%
  filter(model == "All Preds + Intervention")


p<-qplot(preds,betas, facets=outcome*measure~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(all.coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=1, hjust = 1)) 


p

m<- p + scale_y_continuous(limit = c(-1,1))

m
```

## MegaFigure Standardized Score

```{r}
ans.coefs$measure <- "ANS"
ans.coefs$outcome <- "Standarized"
linr2.coefs$measure <- "Linear r^2"
linr2.coefs$outcome <- "Standardized"
ord.coefs$measure <- "Ordinality"
ord.coefs$outcome <- "Standardized"
pae.coefs$measure <- "PAE"
pae.coefs$outcome <- "Standardized"


all.coefs <- bind_rows(ans.coefs, 
                       linr2.coefs, 
                       ord.coefs,
                       pae.coefs
                      ) %>%
  filter(model == "All Preds + Intervention")


p<-qplot(preds,betas, facets=outcome*measure~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(all.coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=.3, hjust = 1)) 


p

#m<- p + scale_y_continuous(limit = c(-.4,.5))

#m
```


## Megafigure, PCA only

```{r}
pca.ans.coefs$measure <- "ANS"
pca.ans.coefs$outcome <- "PCA"
pca.linr2.coefs$measure <- "Linear r^2"
pca.linr2.coefs$outcome <- "PCA"
pca.ord.coefs$measure <- "Ordinality"
pca.ord.coefs$outcome <- "PCA"
pca.pae.coefs$measure <- "PAE"
pca.pae.coefs$outcome <- "PCA"

all.coefs <- bind_rows(
                       pca.ans.coefs,
                       pca.linr2.coefs, 
                       pca.ord.coefs,
                       pca.pae.coefs) %>%
  filter(model == "All Preds + Intervention")


l<-qplot(preds,betas, facets=outcome*measure~year, 
      fill=preds,
      geom="bar",
      stat="identity",
      position="dodge",
      data=subset(all.coefs, preds!="Intercept")) +
  geom_linerange(aes(ymin=betas-se,
                     ymax=betas+se)) +
  geom_hline(yintercept=0, lty=2) + 
  xlab("Predictor") + 
  ylab("Standardized Beta Weight") + 
  scale_colour_discrete(name="Predictor") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.3, hjust = 1)) 


l

#z<- l + scale_y_continuous(limit = c(-1,.5))

#z
```
